---
title: "P8105 HW2"
output: github_document
date: "2022-10-04"
---

```{r}
library(readxl)
library(tidyverse)
library(dplyr)
library(tidyr)

```


# Problem 2

#cleaning Mr. Trash Wheel sheet
```{r}

wheel_data = read_excel("data/Trash Wheel Collection Data.xlsx",
                        range = "A2:N550",
                        sheet = "Mr. Trash Wheel") %>% janitor::clean_names()

wheel_data = wheel_data %>% mutate(sports_balls = as.integer(sports_balls))%>%drop_na(dumpster)%>% mutate(vessel_type="Mr. Trash Wheel")%>%mutate(year=as.double(year))

```
#cleaning Prof Trash Wheel sheet
```{r}
prof_data = read_excel("data/Trash Wheel Collection Data.xlsx",
                        na = "",
                        sheet = "Professor Trash Wheel") %>% janitor::clean_names()%>% drop_na(dumpster)%>%mutate(vessel_type = "Professor Trash Wheel")

```
#combine two data set

```{r}
tidy_data = bind_rows(wheel_data, prof_data)
tidy_data%>% 
  filter(vessel_type == "Professor Trash Wheel")%>%
  select(weight_tons)%>%
  sum()
tidy_data%>%
  filter(year==2020,vessel_type == "Mr. Trash Wheel")%>%
  select(sports_balls)%>%
  sum()

```
After combining these two datasets, the new dataset contains 641 rows and 15 columns, key variables including: dumpster, month, year, date, weight_tons, volume_cubic_yards, plastic_bottles, polystyrene, cigarette_butts, glass_bottles, grocery_bags, chip_bags, sports_balls, homes_powered and the new adding variable vessel_type which helps indicate the data source. For the data we got, we can obtained that the total weight of trash collected by Professor Trash Wheel is 190.12 tons, and the total number of sports balls collected by Mr. Trash Wheel in 2020 is 856.



# Problem 3

#cleaning data in plos-month
```{r}
pols_month = read_csv(file = "data/fivethirtyeight_datasets/pols-month.csv")%>% janitor::clean_names()
tidy_pols = pols_month%>%
  separate(mon,into = c("year", "month", "day"), sep = "-")%>%
  mutate(president = ifelse(prez_dem==0, "gop", "dem"))%>%
  mutate(month = month.name[as.numeric(month)], year = as.integer(year))%>%
  select(-prez_dem, -prez_gop, -day)
  
tidy_pols
  
```

#cleaning data in snp.csv
```{r}
snp = read_csv(file = "data/fivethirtyeight_datasets/snp.csv")%>% janitor::clean_names()
tidy_snp = snp%>%
  separate(date,into = c("month", "day","year"), sep = "/")%>%
  mutate(month = month.name[as.integer(month)], year = as.integer(year), day = as.integer(day))%>%
  mutate(year = as.integer(ifelse(year>=15, year+1900, year+2000)))%>%
  select(year, month, close)%>%arrange(year, month)

tidy_snp
```


#cleaning the unemployment data
```{r}
emp = read_csv(file = "data/fivethirtyeight_datasets/unemployment.csv")%>% janitor::clean_names()
tidy_emp = emp%>%
  pivot_longer(2:13, names_to = "month", values_to = "unemploy")%>%
  mutate(month = recode(month, 
                        'jan'='January',
                        'feb'='February',
                        'mar'='March',
                        'apr'='April',
                        'may'='May',
                        'jun'='June',
                        'jul'='July',
                        'aug'='August',
                        'sep'='September',
                        'oct'='October',
                        'nov'='November',
                        'dec'='December'), year = as.integer(year))

tidy_emp
```

#combine the three datasets
```{r}
tidy_combine = full_join(tidy_pols, tidy_snp)%>% full_join(tidy_emp)

tidy_combine
```
After combing the three data sets, the new dataset obtained contains 645 rows and 12 columns, which gives the dimension of 828 x 11, and the range of years is 68, the key variables are listed as follow:
year
month
gov_gop: # of republican governors with associated time.
sen_gop: # of republican senators with associated time.
rep_gop: # of republican representatives with associated time.
gov_dem: # of democratic governors with associated time.
sen_dem: # of democratic senators with associated time.
rep_dem: # of democratic representatives with associated time.
president: indicate the party of president with associated time.
close: the closing value of stock index.
unemploy_perc: the unemployment percentage with associated time.







